---
title: "Movielens Capstone Project HarvardX"
author: "Cynthia McGinnis"
date: "2/3/2019"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MovieLens Introduction
The MovieLens data set was collected by GroupLens Research. 
Can we predict movie ratings based on user preferance, age of a movie?  Using the MovieLens data set and penalized least squares, the following R script calculates the RMSE based on user ratings, movieId and the age of the movie.

The MovieLens data set contains 10000054 rows, 10677 movies, 797 genres and 69878 users.

The steps performed for analysis of the data 
- Created an age of movie column
- Graphic displays of movie, users and ratings in order to find a pattern or insight to the   
  behavior of the data.
- Explored Genres to determine if ratings could be predicted by genre.
- Explored the Coefficient of Determination R-Squared
- Graphically explored the linear correlation coefficient, r-value
- Calculate RMSE based on movieId, userId, and age of the movie.

After exploring the movies through graphical representations and calculating RMSE, I found the best predictor for ratings was movieId, userId.  The age of the movie didn't change the rmse.

The final RMSE is 0.8252

The following are the libraries I used to explore the data.  Explorations that didn't seem to lead to an insight were taken out of the script.
```{r, include=FALSE}

library(stringr)
library(ggplot2)
library(caret)
library(readr)
library(dplyr)
library(stats)
library(corrplot)
library(tidyr)
library(lubridate)
library(data.table)
library(DT)
library(methods)
library(knitr)
library(grid)
library(gridExtra)
library(Matrix)
library(corrplot)
library(RColorBrewer)
library(corrplot)
library(magrittr)
library(viridis)
library(stringi)
library(matrixStats)
library(heuristica)
library(gam)
library(modelr)
library(tidyr)
library(tidyselect)
library(broom)
library(tibble)
library(purrr)
library(forcats)
library(DBI)
library(hexbin)
```

# DownLoad the data
```{r, echo=TRUE}
dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)
ratings <- read.table(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                      col.names = c("userId", "movieId", "rating", "timestamp"))

```

# Build the data set
```{r}
movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")
movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
                                           title = as.character(title),
                                           genres = as.character(genres))

#Explore the size of the data set
movielens <- left_join(ratings, movies, by = "movieId")
nrow(movielens)
n_distinct(movielens$movieId)
n_distinct(movielens$genres)
n_distinct(movielens$userId)
```

#Validation set will be 10% of the movieLens data
```{r}
set.seed(1)
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]
```
# Make sure userId and movieId in validation set are also in edx set
```{r}
validation <- temp %>% 
  semi_join(edx, by = "movieId") %>%
  semi_join(edx, by = "userId")
```

# Add rows removed from validation set back into edx set
```{r}
removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)
```
##Data Cleaning and, data exploration and Data Visulization
#In order to determine if age of the movie is a factor for predicting rating, I extracted the premier date of the movie, and then calculated the age of the movie.  I will also looked at individual genres for genre effect, as well as, effects of user ratings.
```{r}
head(edx)
glimpse(edx)
```
#How many distinct movie, users and genres
```{r}
n_distinct(edx$movieId)
n_distinct(edx$genres)
n_distinct(edx$userId)
nrow(edx)
```
#Convert Timestamp to year
```{r}
edx <- mutate(edx, year_rated = year(as_datetime(timestamp)))
head(edx)
```
# Extract the premier date and calculate the age of the movie.  Explore whether or not the age of the movie effects predicted ratings
```{r}
#extracting the premier date
premier <- stringi::stri_extract(edx$title, regex = "(\\d{4})", comments = TRUE ) %>% as.numeric()

#Add the premier date
edx_with_title_dates <- edx %>% mutate(premier_date = premier)
head(edx_with_title_dates)
```
#After extracting the premier date from the title, check for accuracy

```{r}
#drop the timestamp
edx_with_title_dates <- edx_with_title_dates %>% select(-timestamp)

head(edx_with_title_dates)

#looking at the dates - are they correct?

edx_with_title_dates %>% filter(premier_date > 2018) %>% group_by(movieId, title, premier_date) %>% summarize(n = n())
edx_with_title_dates %>% filter(premier_date < 1900) %>% group_by(movieId, title, premier_date) %>% summarize(n = n())

#Fix the incorrect dates
edx_with_title_dates[edx_with_title_dates$movieId == "27266", "premier_date"] <- 2004
edx_with_title_dates[edx_with_title_dates$movieId == "671", "premier_date"] <- 1996
edx_with_title_dates[edx_with_title_dates$movieId == "2308", "premier_date"] <- 1973
edx_with_title_dates[edx_with_title_dates$movieId == "4159", "premier_date"] <- 2001
edx_with_title_dates[edx_with_title_dates$movieId == "5310", "premier_date"] <- 1985
edx_with_title_dates[edx_with_title_dates$movieId == "8864", "premier_date"] <- 2004
edx_with_title_dates[edx_with_title_dates$movieId == "1422", "premier_date"] <- 1997
edx_with_title_dates[edx_with_title_dates$movieId == "4311", "premier_date"] <- 1998
edx_with_title_dates[edx_with_title_dates$movieId == "5472", "premier_date"] <- 1972
edx_with_title_dates[edx_with_title_dates$movieId == "6290", "premier_date"] <- 2003
edx_with_title_dates[edx_with_title_dates$movieId == "6645", "premier_date"] <- 1971
edx_with_title_dates[edx_with_title_dates$movieId == "8198", "premier_date"] <- 1960
edx_with_title_dates[edx_with_title_dates$movieId == "8905", "premier_date"] <- 1992
edx_with_title_dates[edx_with_title_dates$movieId == "53953", "premier_date"] <- 2007

```
#Calculate the age of the movie
```{r}
#Calculate the age of a movie
edx_with_title_dates <- edx_with_title_dates %>% mutate(age_of_movie = 2018 - premier_date, 
                                                        rating_date_range = year_rated - premier_date)
head(edx_with_title_dates)
```
# Graph the data
```{r, message=FALSE}
#Distribution of Movie Ratings
edx %>% group_by(movieId) %>% summarize(n = n()) %>%
  ggplot(aes(n)) + geom_histogram(fill = "cadetblue3", color = "grey20", bins = 10) +
  scale_x_log10() +
  ggtitle("Number of Movies Ratings")



```
#Number of Ratings by userId
```{r, message=FALSE}
#Distribution of Users
edx %>% group_by(userId) %>% summarize(n = n()) %>%
  ggplot(aes(n)) + geom_histogram(fill = "cadetblue3", color = "grey20", bins = 10) +
  scale_x_log10() + 
  ggtitle("Number of Users Ratings")

```

# Calculate movie rating average, user rating average, average rating by age of movie, average rating by year
```{r}
#Movie rating averages
movie_avgs <- edx_with_title_dates %>% group_by(movieId) %>% summarize(avg_movie_rating = mean(rating))
user_avgs <- edx_with_title_dates %>% group_by(userId) %>% summarize(avg_user_rating = mean(rating))
year_avgs <- edx_with_title_dates%>% group_by(year_rated) %>% summarize(avg_rating_by_year = mean(rating)) #year the movie was rated
age_avgs <- edx_with_title_dates %>% group_by(age_of_movie) %>% summarize(avg_rating_by_age = mean(rating)) #age of movie
head(age_avgs)

```

```{r}
head(user_avgs)
```


#What is the relationship to the age of a movie and the movies average rating?
#Graph age of movie vs average movie rating

```{r, message=FALSE}
# age of movie vs average movie rating
age_avgs %>%
  ggplot(aes(age_of_movie, avg_rating_by_age)) +
  geom_point() +
  ggtitle("Age of a Movie vs Average Movie Rating")
```
# The above plot shows more variability as movies age.  The plot, also, shows higher ratings the older a movies is up to 90 years old, then the ratings drop.  

```{r, message=FALSE}
# userId vs average movie rating
user_avgs %>%
  ggplot(aes(userId, avg_user_rating)) +
  geom_point(alpha = 1/20, colour = "blue") +
  ggtitle("User vs Average User Rating")
```
#From the above graph, we can see average ratings by user are pretty consistent between 2.5 and 4.5

#Calculating the lm of the age of a movie vs average rating
```{r}
summary(lm(avg_rating_by_age ~ age_of_movie, data = age_avgs))
```
#We can see that R-square is small at 0.30

#Plot the Residuals
```{r, message=FALSE}
avg_rating.lm <- lm(avg_rating_by_age ~ age_of_movie, data = age_avgs)
avg_rating.res <- resid(avg_rating.lm)

plot(age_avgs$age_of_movie, avg_rating.res,
     ylab='Residuals', xlab='age_of_movie', 
     main = 'Average Rating by Age of Movie') + abline(0,0)

```
#The R-squared is fairly small at 0.30; 30% of the variation in movie ratings can be prdicted by 
# explore the data graphically to see if age of the movie and rating are coorelated
```{r, message=FALSE}
#Movies less than 75 years old
age_of_movie_less_than75 <- age_avgs %>% filter(age_of_movie <75)

# age of movie less than 75 years old vs average movie rating
age_of_movie_less_than75 %>%
  ggplot(aes(age_of_movie, avg_rating_by_age)) +
  geom_point() +
  ggtitle("Age of a Movie vs Average Movie Rating")


```
#Calculate the R-squared value
```{r}
age_lessthan75_rating.lm <- lm(avg_rating_by_age ~ age_of_movie, data = age_of_movie_less_than75)
summary(age_lessthan75_rating.lm)

```
#The R-squared increased to 0.745 

#Plot the residuals
```{r, message=FALSE}

head(age_of_movie_less_than75)

age_lessthan75.res <- resid(age_lessthan75_rating.lm)

plot(age_of_movie_less_than75$age_of_movie, age_lessthan75.res,
     ylab='Residuals', xlab='age_of_movie', 
     main = 'Average Rating by Age of Movie') + abline(0,0)
```

#Let's look at moveies between 20 and 75 years old as the graph looks more linear in that time frame
```{r}
#Movies between 20 and 75 years old
age_between20_and_75 <- age_avgs %>% filter((age_of_movie > 20) & (age_of_movie < 75))
```

```{r, message=FALSE}
# graph the age of movie between 30 and 75 years old
age_between20_and_75 %>%
  ggplot(aes(age_of_movie, avg_rating_by_age)) +
  geom_point() + ggtitle("Movies between 30 and 75 years old vs average movie rating")

```
#The plot above appears to be a linear trend; however, the r-square is 0.69
```{r}
summary(lm(avg_rating_by_age ~ age_of_movie, data = age_between20_and_75))
```
#The R-squared value is lower at 0.6981

```{r, message=FALSE}
# graph the age of movie between 20 and 40 years old
age_between20_and_40 <- age_avgs %>% filter((age_of_movie > 20) & (age_of_movie < 40))

age_between20_and_40 %>%
  ggplot(aes(age_of_movie, avg_rating_by_age)) +
  geom_point() + ggtitle('Age of Movie between 20 and 40 years old')

```
#The above graph is displying a linear trend with older movies having higher ratings

```{r}
#calculate a linear model
summary(lm(avg_rating_by_age ~ age_of_movie, data = age_between20_and_40))
```
# The R-squared value is much higher than at 0.71
```{r, message=FALSE}
#Movies between 0 and 30 years old
age_less_than30 <- age_avgs %>% filter((age_of_movie < 30))

#Graph movies less than 30 years old and average movie rating
age_less_than30 %>%
  ggplot(aes(age_of_movie, avg_rating_by_age)) +
  geom_point() + ggtitle('Age of Movie less then 30 years old')
```
#For movies less than 30 years old there appears to be quite a bit of variation.  We can see from the linear model that r-squared is nearly zero.
```{r}
summary(lm(avg_rating_by_age ~ age_of_movie, data = age_less_than30))
```
#The age of a movie did seem to effect the outcome of the average rating.  This is possibly due to a higher number of ratings for older movies.  

#Do Genres have an effect on ratings?
##I extracted the genres from the data with the idea to do an analysis on each genre.  Some of the exploration I did here was removed as it didn't appear to effect the RMSE and this analysis keep growing! But I did get some nice graphs pertaining to genres.
```{r}

#Genres split the data into single genres
dat <- edx_with_title_dates %>% separate_rows(genres, sep ="\\|")

head(dat)

```
#Count the number of movies using movieId in each genre
```{r}
genre_count_by_movieId <- dat %>% group_by(movieId, genres) %>% summarize(n = n())
head(genre_count_by_movieId)
```
#Total number of movies in each genre
```{r}
number_of_genres <- dat %>% group_by(genres) %>% summarize(n = n())
number_of_genres
```
#List the genres.  Movies are either in one genre or multiple genres
```{r}
genre_list <- number_of_genres$genres
genre_list
```

#Explore the distribution of ratings by genre
```{r}
#Distribution of Ratings per Genre
temp <- dat %>%
  group_by(genres) %>%
  summarize(n=n()) %>%
  ungroup() %>%
  mutate(sumN = sum(n), percentage = n/sumN) %>%
  arrange(-percentage)
```
#Bar Graph of Genre's
```{r, message=FALSE}
temp %>%
  ggplot(aes(reorder(genres, percentage), percentage, fill= percentage)) +
  geom_bar(stat = "identity") + coord_flip() +
  scale_fill_distiller(palette = "YlOrRd") + labs(y = "Percentage", x = "Genre") +
  ggtitle("Distribution of Genres by Percent Rated")
```
#From the above graph, we can see Drama had the highest percentage of ratings.

#Genre's Mean rating
```{r, message=FALSE}
temp <- dat %>%
  group_by(genres) %>%
  summarize(mean_rating_by_genre=mean(rating)) %>%
  arrange(-mean_rating_by_genre)

temp %>%
  ggplot(aes(reorder(genres, mean_rating_by_genre), mean_rating_by_genre, fill= mean_rating_by_genre)) +
  geom_bar(stat = "identity") + coord_flip() +
  scale_fill_distiller(palette = "YlOrRd") + labs(y = "Mean Rating", x = "Genre") +
  ggtitle("Average Rating of Genres")
```
#Film Noir had the highest average rating, while Horror had the lowest average rating.

#Explore movie ratings based on number of ratings and value of the rating
```{r}
#Graph of movies with more than 10000 ratings and a mean rating greater than 4. 
avg_rating_greater_than_4 <- edx %>% group_by(title) %>%
     summarize(mean_rating= mean(rating), n = n()) %>% filter(mean_rating >=4) %>% arrange(desc(n, mean_rating))

avg_rating_greater_than_4 %>% filter(n >=10000) %>%
  ggplot(aes(reorder(title, n), n, fill = n)) +
  geom_bar(stat = "identity") + coord_flip() + scale_fill_distiller(palette = "PuBuGn") + xlab("Movie") +ylab('Number of Ratings') +
  ggtitle("Movies with an average rating\ngreater than or equal to 4\nand Number of Ratings  > 10000") 


```
# Examine Movies with ratings between 3 and 4 and more than 10000 ratings
```{r, message=FALSE}
avg_between3_4 <- edx %>% group_by(title) %>%
     summarize(mean_rating= mean(rating), n = n()) %>% filter(n > 10000, (mean_rating >= 3 & mean_rating < 4)) %>% arrange(desc(n, mean_rating))

p <- avg_between3_4 %>% slice(1:40) 
p %>% 
  ggplot(aes(reorder(title, n), n, fill = n)) + 
  geom_bar(stat = "identity") + coord_flip() + scale_fill_distiller(palette = "PuBuGn") +
  ggtitle("Average ratings 3<= r < 4 and n > 10000") + xlab('Movie') + ylab('Number of Ratings') +
  theme_classic()

```

#Movies with an average rating between 2 and 3  lets look at number of ratings greater than 5000
```{r, message=FALSE}
avg_between2_3 <- edx %>% group_by(title) %>%
     summarize(mean_rating= mean(rating), n = n()) %>% filter(n > 5000, (mean_rating >= 2 & mean_rating < 3)) %>% arrange(desc(n, mean_rating))


avg_between2_3 %>% 
  ggplot(aes(reorder(title, n), n, fill = n)) + 
  geom_bar(stat = "identity") + coord_flip() + scale_fill_distiller(palette = "PuBuGn") +
  ggtitle("Average ratings 2<= r < 3 and n > 5000") + xlab('Movie') + ylab('Number of Ratings') +
  theme_classic()

```
#Less than 10000 ratings and a rating less than 2 and number of ratings greater than 500
```{r, message=FALSE}
avg_rating_less_than_2 <- edx %>% group_by(title) %>%
     summarize(mean_rating= mean(rating), n = n()) %>% filter(n > 500, mean_rating < 2) %>% arrange(desc(n, mean_rating))

avg_rating_less_than_2 %>% 
  ggplot(aes(reorder(title, n), n, fill = n)) + 
  geom_bar(stat = "identity") + coord_flip() + scale_fill_distiller(palette = "PuBuGn") +
  ggtitle("Average ratings  < 2") + xlab('Movie') + ylab('Number of Ratings') +
  theme_classic()

```

#Compute the least squares for movieId
```{r, message=FALSE}
#Which movies have a large number of ratings and a rating larger than the average mu
mu <- mean(edx$rating)
edx %>% group_by(title) %>%
  summarize(b_i = mean(rating - mu), n = n()) %>% filter(b_i > 0.5, n > 10000) %>%
  ggplot(aes(reorder(title, b_i), b_i, fill = n)) +
  geom_bar(stat = "identity") + coord_flip() + scale_fill_distiller(palette = "PuBuGn") +
  ggtitle("") + xlab("Movie Title") +
  ggtitle("Movie rating - mu,\nfor Number of ratings > 10000") +
  theme_classic()

```
#Regularized Movie Averages
```{r, message=FALSE}
movie_avgs <- edx %>% group_by(movieId) %>% summarize(b_i = mean(rating - mu))
movie_reg_avgs <- edx %>% 
  group_by(movieId) %>%
  summarize(b_i = sum(rating - mu)/(n()+1), n_i = n())

movie_titles <- edx %>% select(movieId, title) %>% distinct()


edx_with_avgs <- edx %>% group_by(title, movieId) %>% summarize(n = n()) %>%
  left_join(movie_reg_avgs, by = "movieId") %>%
  arrange(desc(b_i, n))

edx_with_avgs %>% filter(n > 15000) %>%
  ggplot(aes(reorder(title, b_i), b_i, fill = n)) +
  geom_bar(stat = "identity") + coord_flip() + scale_fill_distiller(palette = "PuBuGn") +
  ggtitle("") + xlab("Movie Title") + ggtitle('Regularized Movie Averages\nfor Number of Ratings > 20000') +
  theme_classic()


```
#Regularized Movie Averages for the movies with regularized ratings less than 2
```{r, message=FALSE}

head(edx_with_avgs)
p <- edx_with_avgs %>% arrange(b_i) %>% filter(b_i < -2) %>% arrange((b_i))
p

p %>% 
 ggplot(aes(reorder(title, b_i), b_i, fill = n)) +
  geom_bar(stat = "identity") + coord_flip() + scale_fill_distiller(palette = "PuBuGn") +
  ggtitle("") + xlab("Movie Title") + ggtitle('Regularized Movie Averages b_i < -2') +
  theme_classic()
```
#Movies with number of ratings larger than 1000 and regularized average less than 0.
```{r, message=FALSE}
edx_with_avgs %>% filter(n > 10000, b_i < 0.0) %>%
  ggplot(aes(reorder(title, b_i), b_i, fill = n)) +
  geom_bar(stat = "identity") + coord_flip() + scale_fill_distiller(palette = "PuBuGn") +
  ggtitle("") + xlab("Movie Title") + ggtitle('Regularized Movie Averages\nfor Number of Ratings > 20000') +
  theme_classic()

```

#Explore correlation between ratings, users, movieId age of movie and number of ratings

```{r}
#Is there a correlation

#Number of movie ratings per movie
n_movies_ratings <- edx_with_title_dates %>% group_by(movieId) %>% summarize(n = n())

#Average Movie Rating for each movie
avg_movie_rat <- edx_with_title_dates %>% group_by(movieId) %>% summarize(avg_m_r = mean(rating))

#Create correlation data
cor_dat <- edx_with_title_dates %>% select(rating, movieId, userId, year_rated, age_of_movie, rating_date_range, premier_date) %>%
  left_join(n_movies_ratings, by = "movieId") %>%
  left_join(avg_movie_rat, by = 'movieId')
head(cor_dat)
```
#Graph the correlation
```{r, message=FALSE}
temp <- cor_dat %>% select(one_of("rating", "movieId", "userId", "year_rated", "age_of_movie", 
                                               "rating_date_range", "premier_date", "n", "avg_m_r")) %>% as.matrix()
M <- cor(temp, use = "pairwise.complete.obs")

corrplot(M, order = "hclust", addrect = 2, type = "lower", col = brewer.pal(n = 8, name = "RdBu"))
```
#What is the effect of the age of the movie
```{r, message=FALSE}
corr_by_age_of_movie <- cor_dat %>% filter((age_of_movie >20) & (age_of_movie < 70))
temp <- corr_by_age_of_movie %>% select(one_of("rating", "movieId", "userId", "year_rated", "age_of_movie", 
                                  "rating_date_range", "n", "premier_date", "avg_m_r")) %>% as.matrix()
M <- cor(temp, use = "pairwise.complete.obs")

corrplot(M, order = "hclust", addrect = 2, type = "lower", col = brewer.pal(n = 8, name = "RdBu"))
```

```{r, message=FALSE}

#Is there a relationship between number of ratings and the average rating
get_cor <- function(df){
  m <- cor(df$x, df$y, use="pairwise.complete.obs");
  eq <- substitute(italic(r) == cor, list(cor = format(m, digits = 2)))
  as.character(as.expression(eq));
}

#Number of ratings vs avg movie ratings
cor_dat %>%
  ggplot(aes(n, avg_m_r)) + stat_bin_hex(bins = 50) + scale_fill_distiller(palette = "Spectral") +
  stat_smooth(method = "lm", color = "orchid", size = 1) +
  annotate("text", x = 20000, y = 2.5, label = get_cor(data.frame(x = cor_dat$n, y = cor_dat$avg_m_r)), 
           parse = TRUE, color = "orchid", size = 7) + ylab("Average Movie Rating") + xlab("Number of Ratings")
```
#Is there an Age Effect on Movie Ratings?
```{r, message=FALSE}
cor_dat %>% 
  ggplot(aes(age_of_movie, avg_m_r)) + stat_bin_hex(bins = 50) + scale_fill_distiller(palette = "Spectral") +
  stat_smooth(method = "lm", color = "orchid", size = 1) +
  annotate("text", x = 75, y = 0.9, label = get_cor(data.frame(x = corr_by_age_of_movie$age_of_movie, y = corr_by_age_of_movie$avg_m_r)), 
           parse = TRUE, color = "orchid", size = 7) + ylab("Average Movie Rating") + xlab('Age of Movie')
```

#Calculate the RMSE
```{r, message=FALSE}

#RMSE function
RMSE <- function(true_ratings, predicted_ratings){
  sqrt(mean((true_ratings - predicted_ratings)^2))
}

#Choose the tuning value
lambdas <- seq(0,5,.5)
rmses <- sapply(lambdas, function(l){
  mu <- mean(edx_with_title_dates$rating)
  
  b_i <- edx_with_title_dates %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - mu)/(n() + l))
  
  b_u <- edx_with_title_dates %>%
    left_join(b_i, by='movieId') %>% 
    group_by(userId) %>%
    summarize(b_u = sum(rating - b_i - mu)/(n() +l))
  
  predicted_ratings <- edx_with_title_dates %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    mutate(pred = mu + b_i +  b_u) %>% .$pred
  
  return(RMSE(predicted_ratings, edx_with_title_dates$rating))
})

qplot(lambdas, rmses)
lambdas[which.min(rmses)]
```
# Using the model on the Validation data
```{r, message=FALSE}
mu <- mean(validation$rating)
l <- 0.15
b_i <- validation %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - mu)/(n() + l))
  
b_u <- validation %>%
    left_join(b_i, by='movieId') %>% 
    group_by(userId) %>%
    summarize(b_u = sum(rating - b_i - mu)/(n() +l))
  
predicted_ratings <- validation %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    mutate(pred = mu + b_i +  b_u) %>% .$pred

RMSE(predicted_ratings, validation$rating)
```
# I originally calculated a b_a for the age of a movie but found it didn't lower my RMSE so took it out and didn't include it in this script. 
#I used movieId and userId to calculate the RMSE and was able to achieve an RMSE = 0.826

#The code below utilizes the package "Metrics", which resulted in the same RMSE.  I included this as a check for my calculations.
```{r}

library(Metrics)
rmse(validation$rating, predicted_ratings)
```
#This was an amazing assignment; I'm sure I did way more than I needed to as this script is quite long.  However, I wanted to use this data as a learning tool to explore as much as possible from what I've learned in this R adventure.
